Chemistry of fossil Pinus pollen and the effect of acetolysis

### 2021-06-01 manuscript.Rmd
rewrote first results paragraph to include plsr coefficients

PCA on grains for exploration

@@

There are differences between fresh and fossil grains (Fig. \@ref(fig:meanDAL), \@ref(fig:meanMFM) and \@ref(fig:meanTSK).
The prominent and characteristic sporopollenin peaks (1600, 1510 and 1170 cm^-1^) are visible in the fossil pollen, while other parts of the spectra show substantial differences.
Most notibly the lipid peak at 1745 cm^-1^, caused by triglycerides, is not visible in the fossil pollen grains, neither acetolysed or raw.
There are also several differences in the carbohydrate region from 1000 to 1250 cm^-1^ between fresh and fossil pollen grains.
A peak at 1030  and ~1200 cm^-1^ is visible in the acetolysed pollen but not in the fresh spectra, while we see a broad peak at 1090-1100 cm^-1^ in the SPT spectra of all three cores.
This peak is not present in either fresh pollen or acetolysed pollen, additionally Tiefer See shows a similar peak at 1630.

PCA ordination shows that the fossil pollen are clearly separated from the fresh pollen, but not clearly separable into the three cores (Fig \@ref(fig:pcaacetsel)).
Tiefer See pollen is separated from the other two cores, which cluster closer together.
This is the same for acetolysed and SPT pollen. 
A PCA-LDA model to differentiate fossil pollen into the three cores achieved ~60% accuracy to classify MFM and DAL pollen and 85% accuracy to identify TSK pollen (REF TO SM).

>>>>

PLSR second deriv on treatment (fresh vs fossil)

@@

There are differences between modern and fossil grains (Fig. \@ref(fig:meanDAL), \@ref(fig:meanMFM) and \@ref(fig:meanTSK).
The prominent and characteristic sporopollenin peaks (1600, 1510 and 1170 cm^-1^) are visible in the fossil pollen, while other parts of the spectra show substantial differences.
Most notibly the lipid peak at 1745 cm^-1^, caused by triglycerides, is not visible in the fossil pollen grains, neither in the SPT treated or acetolysed pollen.
There are also several differences in the "fingerprint" region (1500 to 1800 cm^-1^) and carbohydrate region from 1000 to 1250 cm^-1^ between modern and fossil pollen grains.
A peak at 1030 and ~1200 cm^-1^ is visible in the acetolysed pollen but not in the modern spectra, while we see a broad peak at 1090-1100 cm^-1^ in the SPT spectra of all three cores.
This peak is not present in either modern pollen or acetolysed pollen, additionally Tiefer See shows a similar peak at 1630.

After visual inspection of the spectra, the chemical differences between modern and fossil polle were explored with a PLSR.
The ordination showed that the fossil pollen are clearly separated from the modern pollen (Fig \@ref(fig:plsrord)).
The regression coefficients reflect our observations from the mean spectra and show high importance of the 1745 lipid peak, the 1500 - 1800 region and the 1170 peak.

<!-- 
Tiefer See pollen is separated from the other two cores, which cluster closer together.
This is the same for acetolysed and SPT pollen. 
A PCA-LDA model to differentiate fossil pollen into the three cores achieved ~60% accuracy to classify MFM and DAL pollen and 85% accuracy to identify TSK pollen (REF TO SM).

-->
